---
title: "Coursera_Week2"
author: "JS"
date: "September 29, 2016"
fig_caption: yes
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='figures/',warning=FALSE, message=FALSE)
```

# Coursera Reproducible Research Week 2


## Loading the data
```{r }
table<-read.csv(file="activity.csv",header = TRUE,colClasses = c("numeric","Date","numeric"))

library(ggplot2)
library(dplyr)

```

## Q1: What is mean total number of steps taken per day? NA omited

Make a histogram of the total number of steps taken each day
 
```{r }
    dailyData <- table %>% group_by(date) %>% summarise(Steps = sum(steps))
    dailyHist <-ggplot(data=dailyData,aes(x=Steps))+geom_histogram()+ggtitle("Total Number of Steps Each Day")
    dailyHist
    
    dailyMean <-mean(dailyData$Steps,na.rm = TRUE)
    dailyMean
    dailyMed <-median(dailyData$Steps,na.rm = TRUE)
    dailyMed
``` 

## Q2: What is the average daily activity pattern?
 
```{r }
    dailyInterval <- table %>% group_by(interval) %>% summarise(avgSteps = mean(steps,na.rm=TRUE))
    dailyLine <- ggplot(data=dailyInterval,aes(x=interval,y=avgSteps))+geom_line()+ggtitle("Average Number of Steps during a Day")
    dailyLine
    print(paste0("On average, the maximum number of steps is during the inverval: ",dailyInterval[dailyInterval$avgSteps==max(dailyInterval$avgSteps),1]))
``` 
  
## Q3: Imputing missing values
  Calculate Nb of NAs
  
```{r }
    summary(table)
    print(paste0("The sum of NA's in the dataset is ",sum(is.na(table))))
```

  Fill NA values by the Average of the interval
```{r }
    AggTable <- merge(table,dailyInterval,by=c("interval"))
    AggTable$FixSteps <- ifelse(is.na(AggTable$steps),AggTable$avgSteps,AggTable$steps)
```

   Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r }
    Table_exNA <- AggTable[,c("date","interval","FixSteps")]
    colnames(Table_exNA)[3]<-"steps"
```

  Histogram of the total number of steps taken each day
```{r }
    dailyData_exNA <-  Table_exNA %>% group_by(date) %>% summarise(Steps = sum(steps))
    dailyHist_exNA <-ggplot(data=dailyData_exNA,aes(x=Steps))+geom_histogram()+ggtitle("Total Number of Steps Each Day - ex NA")
    dailyHist_exNA
```

  Calculate Mean and Median
```{r }
    dailyMean_exNA <-mean(dailyData_exNA$Steps)
    dailyMean_exNA
    dailyMed_exNA <- median(dailyData_exNA$Steps)
    dailyMed_exNA
```
  Both mean and median are really close. this is logic as we replace NA values by the average
  
    
## Q4: Are there differences in activity patterns between weekdays and weekends?
    
    Define Weekday factor
```{r }
  
    Table_exNA$WeekdayType <- ifelse(weekdays(Table_exNA$date) %in% c("Saturday","Sunday"),"weekend","weekday")
    
    WeekDayInterval <- Table_exNA %>% group_by(interval,WeekdayType) %>% summarise(avgSteps = mean(steps,na.rm=TRUE))
    WeekDayPlot <- ggplot(data=WeekDayInterval,aes(x=interval,y=avgSteps))+geom_line()+facet_grid(WeekdayType ~ .)
    WeekDayPlot
```
    
    